---
title: "HW 6: Pronto!"
author: "Terrence Pope"
date: "May 9, 2018"
output:
  html_document:
    toc: true
    toc_float: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=TRUE, message=TRUE)
library(tidyverse)
library(magrittr)
library(lubridate)
```
# Instructions

> Pronto! is Seattle's bike sharing program, which launched in fall 2014. You may have seen the green bike docks around campus. It was also [in the news frequently](http://www.king5.com/news/local/seattle/city-launches-investigation-into-sdot-director-over-pronto-ties/97249954) and [eventually shut down](http://www.seattletimes.com/seattle-news/transportation/seattle-pronto-bike-share-shutting-down-friday/).

> You will be using data from the [2015 Pronto Cycle Share Data Challenge](https://www.prontocycleshare.com/datachallenge). These are available for download as a 75 MB ZIP file from <https://s3.amazonaws.com/pronto-data/open_data_year_one.zip>. Once unzipped, the folder containing all the files is around 900 MB. The `open_data_year_one` folder contains a `README.txt` file that you should reference for documentation.

> Questions for you to answer are as quoted blocks of text. Put your code used to address these questions and any comments you have below each block. Remember the guiding DRY principle: **Don't Repeat Yourself!**


# PART 1

## Loading the Data

> Set your working directory to be the `open_data_year_one` folder. Then use the `list.files()` command to return a character vector giving all the files in that folder, and store it to an object called `files_in_year_one`. Then use vector subsetting on `files_in_year_one` to remove the entries for `README.txt` (which isn't data) and for `2015_status_data.csv` (which is massive and doesn't have interesting information, so we're going to exclude it). Thus, `files_in_year_one` should be a **character vector with three entries**.

```{r loading}
setwd("~/Desktop/UW/Spring '18/Intro to R/Week 7/Assignment")

files_in_year_one <- list.files("open_data_year_one")
  
files_in_year_one <- files_in_year_one[-2][-4]
```

> We want to read the remaining .csv files into data frames stored in a **list** called `data_list`. Preallocate this using `data_list <- vector("list", length(files_in_year_one))`.

```{r vector}
data_list <- vector("list", length(files_in_year_one))
```

> We would like the names of the list entries to be simpler than the file names. For example, we want to read the `2015_station_data.csv` file into `data_list[["station_data"]]`, and `2015_trip_data.csv` into `data_list[["trip_data"]]`. So, you should make a new vector called `data_list_names` giving the names of the objects to read in these CSV files to using `files_in_year_one`. Use the `str_sub()` function in `library(stringr)` to keep the portion of the `files_in_year_one` entries starting from the sixth character (which will drop the `2015_` part) and stopping at number of characters of each filename string, minus 4 (which will drop the `.csv` part). Remember to load `stringr` with `library` and use `?str_sub` in the console to get help on the function. 

```{r nameset}
data_list_names <- str_sub(files_in_year_one, 6, -5)
#data_list[[1]] <- read.csv("2015_station_data.csv")
#data_list[[2]] <- read.csv("2015_trip_data.csv")
#data_list[[3]] <- read.csv("2015_weather_data.csv")
#data_list
```

> Set the names for `data_list` using the `names()` function and the `data_list_names` vector.

```{r naming}
names(data_list) <- data_list_names
```

> Then, write a `for()` loop that uses `read_csv()` from the `readr` package to read in all the CSV files contained in the ZIP file, `seq_along`ing the `files_in_year_one` vector. Store each of these files to its corresponding entry in `data_list`. The [data download demo](https://rebeccaferrell.github.io/CSSS508/Lectures/data_download_demo.html) might be a helpful reference.

> You will want to use the `cache=TRUE` chunk option for this chunk --- otherwise you'll have to wait for the data to get read in every single time you knit. You will also want to make sure you are using `read_csv()` in the `readr` package and not base R's `read.csv()` as `readr`'s version is much faster, gives you a progress bar, and won't convert all character variables to factors automatically.

```{r loop, cache=TRUE}
for(i in seq_along(files_in_year_one)) {
    data_list[[i]] <- read_csv(paste0("open_data_year_one/", files_in_year_one[i]))
}
```

## Fixing Data Types

> Run `str()` on `data_list` and look at how the variables came in from using `read_csv()`. Most should be okay, but some of the dates and times may be stored as character rather than dates or `POSIXct` date-time values. We also have lots of missing values for `gender` in the trip data because users who are not annual members do not report gender. 

> First, patch up the missing values for `gender` in `data_list[["trip_data"]]`: if a user is a `Short-Term Pass Holder`, then put `"Unknown"` as their `gender`. Don't make new objects, but rather modify the entries in `data_list` directly (e.g. `data_list[["trip_data"]] <- data_list[["trip_data"]] %>% mutate(...)`.

```{r fix}
str(data_list)
data_list[["trip_data"]] <- data_list[["trip_data"]] %>%
  mutate(gender = ifelse(gender == "Short-Term Pass Holder", "Unknown", gender))
```

> Now, use `dplyr`'s `mutate_at()` and/or `mutate()`, functions from the `lubridate` package, and the `factor` function to (1) fix any date/times, as well as to (2) convert the `usertype` and `gender` variables to factor variables from the trip data. *Don't make new objects*, but rather modify the entries in `data_list` directly.

```{r mutate}
data_list[["trip_data"]] <- data_list[["trip_data"]] %>%
  mutate_at(vars(starttime, stoptime), funs(mdy_hm))

data_list[["trip_data"]] <- data_list[["trip_data"]] %>%
  mutate_at(vars(usertype, gender), funs(factor))

data_list[["station_data"]] <- data_list[["station_data"]] %>%
  mutate_at(vars(online), funs(mdy))

data_list[["weather_data"]] <- data_list[["weather_data"]] %>%
  mutate_at(vars(Date), funs(mdy))
```